### 2020-05-31 paper/paper.tex
Write some of the paper.
- Intro written
  - Background
  - System advantages
  - Use cases
- There's a 'make spell'

\documentclass[11pt,english,twocolumn]{article}

\title{Append-only Datastore}
\author{
	Mingyan Zhao
	\and
	Steven Tung
	\and
	Kevin Krakauer
}
\date{}

\begin{document}

\maketitle

% Calling our operation "put" is inherently confusing, I'm using "append".

\section*{}
We built an eventually consistent, append-only datastore focused on low
read/write latency and high availability. In the common case, clients interact
only with nearby nodes for extremely low latency. We do not provide overwrite
and delete operations. This simple interface guarantees properties making client
implementation simple: the relative order of data for a given key is consistent
and each client communicates only with a single node.

The datastore is eventually consistent. Data is stored in memory for low latency
and in local disk for safety. Nodes can operate when disconnected from the
system, preserving availability in the face of total and long-term partitioning.
We believe this system will be useful in chat and distributed logging
applications.

\section{Introduction}
As organizations increasingly move to the cloud, applications are designed from
the ground up with a distributed architecture (in some cases referred to as
\textit{microservices}). Despite numerous advantages, distributed
application design requires addressing latency and partitioning concerns that
monolithic applications do not have (or in the case of partitioning, are not
solvable).

Our \textit{append-only datastore} addresses latency and partitioning concerns
for append-only workloads. By explicitly distinguishing between a
\textit{master} node and \textit{follower} nodes, we gain several advantages
over other eventually consistent systems:

\begin{enumerate}
	\item Clients communicate only with their nearest follower for extremely
		low latency.
	\item Followers provide linearizability for client appends on a single
		key. That is, all clients writing to the same key of the same
		follower will see linearized writes.
	\item Eventual consistency is minimally disruptive to clients. A client
		may read the data for key $k$ and receive data consisting of 2
		writes: $w_1, w_2$. Write $w_x$ sent through another follower
		preserves the ordering of $w_1$ and $w_2$, so the system
		eventually returns $w_x, w_1, w_2$, $w_1, w_x, w_2$, or $w_1,
		w_2, w_x$ when read.  Because of (2), writes directly to the
		same follower are linearizable.
\end{enumerate}

We also gain the advantages of some more traditionally eventually consistent
systems, such as great partition tolerance \cite{dynamo}. Together, these
properties are highly desirable for many append-only applications. Consider the
following:

\begin{itemize}
	\item \textbf{Chat} - Chat applications are inherently append-only. Users
		in different parts of the globe expect near-instantaneous (only
		a few seconds) latency. Users in the same building, however,
		expect to receive messages instantaneously. Think of a global
		company: teammates send messages and links to nearby teammates,
		expecting them to arrive immediately. Consider a chatroom
		occupied by a team in North American and a team in Asia. Each
		team member sees their local peers' messages immediately and in
		exactly the order that they are sent. It's acceptable for the a
		message from the other team to take a bit longer and appear
		slightly out of order in the logged chat.
	\item \textbf{Distributed logging} - Consider a distributed application
		running in multiple data centers on multiple continents. The
		application is monitored in real time in each datacenter. Logs
		are write-only, and critical for monitoring. By using the
		append-only datastore for logging, the monitoring service:
		\begin{itemize}
			\item Continues operating when a datacenter is
				partitioned from the rest of the system.
			\item Receives logs with extremely low latency from the
				nearby monitored application.
			\item Has an eventually consistent log that can be
				stored for later analysis and debugging.
		\end{itemize}
	% \item \textbf{Distributed data processing}
\end{itemize}

\section{Design}
We made it go super fast.

\subsection{Leader}
% Reference GFS and not transporting the data itself.
It synchronizes stuff.

\subsection{Follower}
% Be sure to discuss any synchronization that happens here, as the follower is
% responsible for synchronizing its the writes to it.
It also sorta synchronizes, but mostly it follows.

\subsection{Clients}
% Describe the API and guarantees exposed to clients.
We have very distinguished clientele.

\section{Implementation}

\section{Evaluation}
We tested using some super cool gcloud stuff.

\subsection{Workload 1}
Oh wow, we're so fast.

\subsection{Workload 2}
Oh no, we're unexpectedly not that fast.

\section{Related Work}
% Talk about Kafka
This is the most filler of all filler sections. Look, a reference! \cite{dynamo}

\section{Future Work}
Make it actually work.

\section{Conclusions}
I'd just like everyone to pat themselves on the back here.

\bibliography{paper} 
\bibliographystyle{ieeetr}

\end{document}