### 2020-06-01 paper/paper.tex
Did a bunch of the follower

\documentclass[11pt,english,twocolumn]{article}

\title{Append-only Datastore}
\author{
	Mingyan Zhao
	\and
	Steven Tung
	\and
	Kevin Krakauer
}
\date{}

\begin{document}

\maketitle

% Calling our operation "put" is inherently confusing, I'm using "append".

\section*{}
We built an eventually consistent, append-only datastore focused on low
read/write latency and high availability. In the common case, clients interact
only with nearby nodes for extremely low latency. We do not provide overwrite
and delete operations. This simple interface guarantees properties making client
implementation simple: the relative order of data for a given key is consistent
and each client communicates only with a single node.

The datastore is eventually consistent. Data is stored in memory for low latency
and in local disk for safety. Nodes can operate when disconnected from the
system, preserving availability in the face of total and long-term partitioning.
We believe this system will be useful in chat and distributed logging
applications.

\section{Introduction}
As organizations increasingly move to the cloud, applications are designed from
the ground up with a distributed architecture (in some cases referred to as
\textit{microservices}). Despite numerous advantages, distributed
application design requires addressing latency and partitioning concerns that
monolithic applications do not have (or in the case of partitioning, are not
solvable).

Our \textit{append-only datastore} addresses latency and partitioning concerns
for append-only workloads. It is designed to run as a service distributed
globally across multiple data centers. By explicitly distinguishing between a
\textit{leader} node and \textit{follower} nodes, we gain several advantages
over other eventually consistent systems:

\begin{enumerate}
	\item Clients communicate only with their nearest follower for extremely
		low latency.
	\item Followers provide linearizability for client appends on a single
		key. That is, all clients writing to the same key of the same
		follower will see linearized writes.
	\item Eventual consistency is minimally disruptive to clients. A client
		may read the data for key $k$ and receive data consisting of 2
		writes: $w_1, w_2$. Write $w_x$ sent through another follower
		preserves the ordering of $w_1$ and $w_2$, so the system
		eventually returns $w_x, w_1, w_2$, $w_1, w_x, w_2$, or $w_1,
		w_2, w_x$ when read.  Because of (2), writes directly to the
		same follower are linearizable.
\end{enumerate}

We also gain the advantages of some more traditionally eventually consistent
systems, such as great partition tolerance \cite{dynamo}. Together, these
properties are highly desirable for many append-only applications. Consider the
following:

\begin{itemize}
	\item \textbf{Chat} - Chat applications are inherently append-only. Users
		in different parts of the globe expect near-instantaneous (only
		a few seconds) latency. Users in the same building, however,
		expect to receive messages instantaneously. Think of a global
		company: teammates send messages and links to nearby teammates,
		expecting them to arrive immediately. Consider a chatroom
		occupied by a team in North American and a team in Asia. Each
		team member sees their local peers' messages immediately and in
		exactly the order that they are sent. It's acceptable for the a
		message from the other team to take a bit longer and appear
		slightly out of order in the logged chat.
	\item \textbf{Distributed logging} - Consider a distributed application
		running in multiple data centers on multiple continents. The
		application is monitored in real time in each datacenter. Logs
		are write-only, and critical for monitoring. By using the
		append-only datastore for logging, the monitoring service:
		\begin{itemize}
			\item Continues operating when a datacenter is
				partitioned from the rest of the system.
			\item Receives logs with extremely low latency from the
				nearby monitored application.
			\item Has an eventually consistent log that can be
				stored for later analysis and debugging.
		\end{itemize}
	% \item \textbf{Distributed data processing}
\end{itemize}

\section{Design}
% TODO: Add architecture diagram.
Nodes in the append-only datastore are classified in a simple hierarchy as seen
in TODO(add the architecture diagram). A single \textit{leader} directs multiple
\textit{followers}. Clients connnect to and communicate with a nearby follower,
likely running in the same datacenter, to minimize latency. Our design tolerates
arbitary partitions between nodes and ensures an eventually consistent view of
data.

Clients are exposed the following API by followers:
\begin{itemize}
	\item \texttt{append(key, data)} - Appends \texttt{data} to the existing
		data for \texttt{key}. Guarantees that \texttt{data} appears
		after any data that the client already read for \texttt{key}.
	\item \texttt{get(key)} - Gets the data for \texttt{key}, expressed as a
		list of indexed values. For example, the returned data will be
		the result of writes 1, 3, and 5, expressed as $\{w_1, w_3,
		w_5\}$. Subsequent \texttt{get} requests will always return
		$w_1$, $w_3$, and $w_5$ in the same relative order, but other
		writes may be prepended, appended, or interspersed.
\end{itemize}

\subsection{Follower}
% Explain what happens when this is partitioned.
% Explain read-your-own-writes consistency.
% Be sure to discuss any synchronization that happens here, as the follower is
% responsible for synchronizing its the writes to it.
The follower is our most complex component. Its responsibilities include:

\begin{itemize}
	\item Handing \texttt{append} and \texttt{get} requests from clients.
	\item Ordering appends.
	\item Updating the master about appended data.
	\item Receiving the eventual, global ordering of data from the master
		and imposing it on data.
	\item Syncing with other followers to communicate updates and orderings.
\end{itemize}

% TODO: We don't actually do the update-on-get thing yet.
Each follower stores the entire datastore locally in memory. When clients issue
\texttt{get(k)} requests, these are served from the in-memory datastore to minimize
response latency. It also causes the follower to ask the leader for the most
up-to-date data for \texttt{k}, ensuring that frequently accessed data is kept
relatively fresh.

\texttt{append} requests are also fast, limited by the speed of a local disk
write. When a client appends data to key \texttt{k}, the follower writes that
data - along with its index - to local disk before returning to the client. This
ensures that data is not lost upon crashing and rebooting. For increased fault
tolerance, the append should be written to multiple disks locally or to a remote
disk.

As soon as the append is written to disk, the follower (concurrently with
replying to the client) sends an \texttt{update} request to the leader
containing the key being appended to and the index that the follower has
assigned the appended data. The leader resolves and orders the append with any
other concurrent appends from different followers (see next section), and
responds.

There are two types of responses. The first - and likely most common - case
is a simple \texttt{SUCCESS} response. This indicates that the append committed
successfully and the follower has up-to-date data for that key. The second is a
\texttt{NEEDS\_UPDATE} response. This indicates that the data was appended, but
another \texttt{update} preempted it. In this scenario, the follower changes the
locally stored index of the appended data, and sends a \texttt{sync} request
another follower indicated by the leader's response.

Followers use the \texttt{sync} request to get data they are missing,
effectively ``patching the holes" in their local datastore. This ensures the
eventually consistent property of the system.

In the common case, where updates do not overlap...

When concurrent updates do occur...

Both \texttt{update} and \texttt{sync} requests are retried in case of network
failure. If a follower reboots during a \texttt{update}, it can inspect the disk
at boot and see that it has appended data locally that has not been confirmed by
the leader. If a follower reboots during a \texttt{sync}, it will have stale
data until a \texttt{get} or \texttt{append} causes it to update. This keeps the
system simple while retaining its consistency properties.

\subsection{Leader}
% Explain what happens when this is partitioned.
% Reference GFS and not transporting the data itself.
It synchronizes stuff.

\subsection{Clients}
% Describe the API and guarantees exposed to clients.
We have very distinguished clientele.

\subsection{Common Case}

\subsection{Corner Cases}
% Multiple simultaneous writes from c1 and c2 to f1. How do we ensure that the
% GL does not order w2 before w1?

\section{Implementation}

\section{Evaluation}
We tested using some super cool gcloud stuff.

\subsection{Workload 1}
Oh wow, we're so fast.

\subsection{Workload 2}
Oh no, we're unexpectedly not that fast.

\section{Related Work}
% Talk about Kafka
This is the most filler of all filler sections. Look, a reference! \cite{dynamo}

\section{Future Work}
% Use a raft group instead of a single node for each follower and GL.
% Client fallback to other followers.
Make it actually work.

\section{Conclusions}
I'd just like everyone to pat themselves on the back here.

\bibliography{paper} 
\bibliographystyle{ieeetr}

\end{document}